---
title: "CGM in CF - Beta cell function outcomes"
author: "Laura Pyle"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(Hmisc)
library(knitr)
library(broom)

# read in OGTT data
source("C:\\Users\\pylell\\Documents\\GitHub\\Chan-ADA-2019\\CGMInHealthyControls_R_2018-12-21_1936.r")
#data <- data[c("subject_id","min_30_insulin","min_0_insulin","min_30_glucose","lab_ogtt_fasting",
#               "min_30_c_peptide","min_0_c_peptide")]

source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\temp_table1.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\01_functions.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\laura_functions.r")
source("C:\\Users\\pylell\\Documents\\GitHub\\General-code\\iAUC.r")

# calculate beta cell function outcomes
data$isi <- (data$min_30_insulin - data$min_0_insulin) / (data$min_30_glucose-data$lab_ogtt_fasting)
data$isi_cpep <- (data$min_30_c_peptide - data$min_0_c_peptide) / (data$min_30_glucose-data$lab_ogtt_fasting)
data$odi <- (1/data$min_0_insulin) * data$isi
data$odi_cpep <- (1/data$min_0_c_peptide) * data$isi_cpep

# also need the following
# iAUCg, iAUCins/iAUCg, iAUCcpep/iAUCg, iAUCglucagon
#glucose
cols_gluc = c("min_minus_10_glucose","lab_ogtt_fasting",
         "min_10_glucose","min_20_glucose","min_30_glucose",
         "lab_ogtt_1_hour_glucose","min_90_glucose",
         "lab_ogtt_2_hour_glucose","min_150_glucose",
         "min_180_glucose")
ogttdat <- data[,c(cols_gluc,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_gluc]))>=4),]
ogttdat$iaucbase_glucose <- iAUC_glucose(ogttdat,ogttdat$lab_ogtt_fasting,type = "linear")
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

#insulin
cols_ins = c("min_minus_10_insulin","min_0_insulin","min_10_insulin","min_20_insulin","min_30_insulin","min_60_insulin",
             "min_90_insulin","min_120_insulin","min_150_insulin","min_180_insulin")
ogttdat <- data[,c(cols_ins,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_ins]))>=4),]
ogttdat$iaucbase_insulin <- iAUC_insulin(ogttdat,ogttdat$min_0_insulin,type = "linear")
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

#C-peptide
cols_cpep = c("min_minus_10_c_peptide","min_0_c_peptide","min_10_c_peptide","min_20_c_peptide","min_30_c_peptide","min_60_c_peptide",
             "min_90_c_peptide","min_120_c_peptide","min_150_c_peptide","min_180_c_peptide")
ogttdat <- data[,c(cols_cpep,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_cpep]))>=4),]
ogttdat$iaucbase_cpep <- iAUC_cpep(ogttdat,ogttdat$min_0_c_peptide,type = "linear")
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

# glucagon
cols_glucagon = c("min_minus_10_glucagon","min_0_glucagon","min_10_glucagon","min_20_glucagon","min_30_glucagon","min_60_glucagon",
              "min_90_glucagon","min_120_glucagon","min_150_glucagon","min_180_glucagon")
ogttdat <- data[,c(cols_glucagon,"subject_id")]
ogttdat <- ogttdat[which(rowSums(!is.na(ogttdat[,cols_glucagon]))>=4),]
ogttdat <- ogttdat[which(!is.na(ogttdat[,"min_0_glucagon"])),]
ogttdat$iaucbase_glucagon <- iAUC_glucagon(ogttdat,ogttdat$min_0_glucagon,type = "linear")
data <- merge(data,ogttdat,by="subject_id",all.x=TRUE,all.y=TRUE)

data$iAUCins_iAUC_gluc <- data$iaucbase_insulin/data$iaucbase_glucose
data$iAUCcpep_iAUCgluc <- data$iaucbase_cpep/data$iaucbase_glucose

# read in other variables from primary outcome analysis
alldata <- read.csv(file = "H:\\Endocrinology\\Chan\\Chan CF CGM\\Repeat primary analysis\\Data\\export_tim.csv")

# merge datasets - first just figure out who is in which datasets
a <- data[c("subject_id")]
a$inogtt <- rep(1,nrow(a))
b <- alldata[c("subject_id","group")]
b$inprimary <- rep(1,nrow(b))
ab <- merge(a,b,by="subject_id",all=TRUE)
ab <- ab[ab$inogtt==1 & ab$inprimary==1,]
ab <- ab[order(ab$subject_id),] 

# merge full datasets
data$inogtt <- rep(1,nrow(data))
alldata$inredcap <- rep(1,nrow(alldata))
alldata <- merge(alldata,data,by="subject_id")
# keep only those who have non-missing odi
alldata <- alldata[which(!is.na(alldata$odi) & alldata$odi != Inf),]
#View(alldata[c("subject_id","group","min_30_insulin","min_0_insulin","min_30_glucose","lab_ogtt_fasting","min_30_c_peptide","min#_0_c_peptide","isi","isi_cpep","odi","odi_cpep")])

ogttvars <- c("isi","isi_cpep","odi","odi_cpep","iaucbase_glucose","iaucbase_insulin","iaucbase_cpep","iaucbase_glucagon","iAUCins_iAUC_gluc","iAUCcpep_iAUCgluc")
cgmvars <- c("min_sensor","max_sensor","average_sensor","avg_auc","excursions_200","excursions_140_day",
             "excursions_200_day","excursions_over_140","percent_time_60","percent_time_70",
             "percent_time_spent_140","time_spent_60","time_spent_70","time_spent_200","time_spent_min_140",
             "time_spent_min_200","total_auc","standard_deviation","mage")
tab1vars <- c("age", "race_ethnicity","gender","a1c","g_tube_feeds","fev1_percent","fvc","cf_pancreatic","tanner","first_bmiz")

# make factors
alldata$gender <- as.factor(alldata$gender)
alldata$g_tube_feeds <- as.factor(alldata$g_tube_feeds)
alldata$cf_pancreatic <- as.factor(alldata$cf_pancreatic)

# labels
label(alldata$age)="Age"
label(alldata$race_ethnicity)="Race/ethnicity"
label(alldata$gender)="Gender"
label(alldata$a1c)="HbA1c"
label(alldata$g_tube_feeds)="G-tube feeds"
label(alldata$fev1_percent)="FEV-1 percent"
label(alldata$fvc)="FVC"
label(alldata$cf_pancreatic)="Pancreatic insufficiency"
label(alldata$tanner)="Tanner stage"
label(alldata$first_bmiz)="First BMI-Z"
label(alldata$isi)="ISI"
label(alldata$isi_cpep)="ISI C-peptide"
label(alldata$odi)="oDI"
label(alldata$odi_cpep)="oDI C-peptide"
label(alldata$min_sensor)="Min sensor glucose"
label(alldata$max_sensor)="Max sensor glucose"
label(alldata$average_sensor)="Average sensor glucose"
label(alldata$avg_auc)="Average AUC"
label(alldata$excursions_200)="Excursions>200"
label(alldata$excursions_140_day)="Excursions>140/day"
label(alldata$excursions_200_day)="Excursions>200/day"
label(alldata$excursions_over_140)="Excursions>140"
label(alldata$percent_time_60)="% time <60"
label(alldata$percent_time_70)="% time <70"
label(alldata$percent_time_spent_140)="% time >140"
label(alldata$time_spent_60)="Time spent <60"
label(alldata$time_spent_70)="Time spent <70"
label(alldata$time_spent_200)="Time spent >200"
label(alldata$time_spent_min_140)="Time spent >140"
label(alldata$time_spent_min_200)="Time spent >200"
label(alldata$total_auc)="Total AUC"
label(alldata$standard_deviation)="SD"
label(alldata$mage)="MAGE"
label(alldata$iaucbase_glucose)="iAUC glucose"
label(alldata$iaucbase_insulin)="iAUC insulin"
label(alldata$iaucbase_cpep)="iAUC C-peptide"
label(alldata$iaucbase_glucagon)="iAUC glucagon"
label(alldata$iAUCins_iAUC_gluc)="iAUC insulin/iAUC glucose"
label(alldata$iAUCcpep_iAUCgluc)="iAUC C-peptide/iAUC glucose"

npar <- nrow(alldata)

# ANALYSIS

# table 1
tab1 <- final_table(data=alldata,variables=tab1vars,group=alldata$group,margin=2,single=0,ron=2)

# pairwise comparisons for table 1 - HbA1c
a1 <- aov(alldata$a1c~alldata$group)
a1t <- TukeyHSD(a1)
a1t <- tidy(a1t)
a1t <- a1t[,-1]

# compare groups
tab2 <- final_table(data=alldata,variables=c(ogttvars,cgmvars),group=alldata$group,margin=2,single=0,ron=2)

# pairwise comparisons for table 2 - ISI
a2 <- aov(alldata$isi~alldata$group)
a2t <- TukeyHSD(a2)
a2t <- tidy(a2t)
a2t <- a2t[,-1]

# pairwise comparisons for table 2 - ISI C-peptide
a3 <- aov(alldata$isi_cpep~alldata$group)
a3t <- tidy(TukeyHSD(a3))
a3t <- a3t[,-1]

# pairwise comparisons for table 2 - oDI
a4 <- aov(alldata$odi~alldata$group)
a4t <- tidy(TukeyHSD(a4))
a4t <- a4t[,-1]

# pairwise comparisons for table 2 - oDI C-pep
a5 <- aov(alldata$odi~alldata$group)
a5t <- tidy(TukeyHSD(a5))
a5t <- a5t[,-1]

# pairwise comparisons for table 2 - iAUC glucose
a6 <- aov(alldata$iaucbase_glucose~alldata$group)
a6t <- tidy(TukeyHSD(a6))
a6t <- a6t[,-1]

# pairwise comparisons for table 2 - iAUC insulin/iAUC glucose
a7 <- aov(alldata$iAUCins_iAUC_gluc~alldata$group)
a7t <- tidy(TukeyHSD(a7))
a7t <- a7t[,-1]

# pairwise comparisons for table 2 - iAUC C-peptide/iAUC glucose
a8 <- aov(alldata$iAUCcpep_iAUCgluc~alldata$group)
a8t <- tidy(TukeyHSD(a8))
a8t <- a8t[,-1]

# pairwise comparisons for table 2 - max sensor glucose
a9 <- aov(alldata$max_sensor~alldata$group)
a9t <- tidy(TukeyHSD(a9))
a9t <- a9t[,-1]

# pairwise comparisons for table 2 - excursions >140/day
a10 <- aov(alldata$excursions_140_day~alldata$group)
a10t <- tidy(TukeyHSD(a10))
a10t <- a10t[,-1]

# pairwise comparisons for table 2 - excursions >140
a11 <- aov(alldata$excursions_over_140~alldata$group)
a11t <- tidy(TukeyHSD(a11))
a11t <- a11t[,-1]

# pairwise comparisons for table 2 - time spent >140
a12 <- aov(alldata$time_spent_min_140~alldata$group)
a12t <- tidy(TukeyHSD(a12))
a12t <- a12t[,-1]

# pairwise comparisons for table 2 - SD
a13 <- aov(alldata$standard_deviation ~alldata$group)
a13t <- tidy(TukeyHSD(a13))
a13t <- a13t[,-1]

# pairwise comparisons for table 2 - MAGE
a14 <- aov(alldata$mage ~alldata$group)
a14t <- tidy(TukeyHSD(a14))
a14t <- a14t[,-1]

# correlations
corr <- rcorr(as.matrix(alldata[, c(ogttvars,cgmvars)]), type = "pearson")
corr <- flattenCorrMatrix(corr$r, corr$P)
corr <- corr[corr$row %in% ogttvars & corr$column %in% cgmvars,]



```
# Background

The purpose of this analysis is to examine associations between beta cell function and CGM variables in healthy controls and participants with cystic fibrosis with normal glucose tolerance, impaired glucose tolerance, and CFRD.

# Methods

Participants who had both CGM data and OGTT data with 0, 30, and 120 minute timepoints available were included in this analysis. Beta cell function outcomes were calculated as follows:

1) $ISI = (Insulin_{30} - Insulin_{0}) / (Glucose_{30} - Glucose_{0})$
2) $ISI_{Cpep} = (Cpeptide_{30} - Cpeptide_{0}) / (Glucose_{30} - Glucose_{0})$
3) $oDI = (1/Insulin_{0}) * ISI$
4) $oDI_{Cpep} = (1/Cpeptide_{0}) * ISI_{Cpep}$

Integrated AUC (iAUC) for glucose, insulin, C-peptide, and glucagon were calculated as the AUC with respect to the fasting value, using the trapezoidal method.

The distributions of all variables were examined prior to analysis.  Groups were compared using ANOVA or the Wilcoxon test for continuous variables, and the chi-square or Fisher's exact test for categorical variables.  Pearson's correlation coefficient was used to test the association between beta cell function outcomes and CGM outcomes.

# Results

There were `r npar` participants included in this analysis.  Table 1 shows baseline descriptive statistics by treatment group.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab1,caption="Table 1.  Descriptive statistics by treatment group.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.") 
```
\

Pairwise comparisons for HbA1c are shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a1t,caption="Pairwise comparisons for HbA1c using Tukey's HSD test.")
```

Descriptive statistics for beta cell function outcomes and CGM outcomes are shown in Table 2.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(tab2,caption="Table 2.  Descriptive statistics for beta cell function and CGM outcomes by treatment group.  Statistics given are N(%), mean±SD, or median (25th %ile, 75th %ile).  * denotes that the Wilcoxon test was used, and ** denotes that Fisher's exact test was used.")
```
\

Pairwise comparisons for significant results in Table 2 are shown below.

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a2t,caption="Pairwise comparisons for ISI using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a3t,caption="Pairwise comparisons for ISI C-peptide using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a4t,caption="Pairwise comparisons for oDI using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a5t,caption="Pairwise comparisons for oDI C-peptide using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a6t,caption="Pairwise comparisons for iAUC glucose using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a7t,caption="Pairwise comparisons for iAUC insulin/iAUC glucose using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a8t,caption="Pairwise comparisons for iAUC C-peptide/iAUC glucose using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a9t,caption="Pairwise comparisons for max sensor glucose using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a10t,caption="Pairwise comparisons for excursions>140/day using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a11t,caption="Pairwise comparisons for excursions>140 using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a12t,caption="Pairwise comparisons for time spent >140 using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a13t,caption="Pairwise comparisons for SD using Tukey's HSD test.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(a14t,caption="Pairwise comparisons for MAGE using Tukey's HSD test.")
```
\

Pearson correlation coefficients between beta cell function outcomes and CGM outcomes are shown in Table 2.
```{r, results='asis',tidy=TRUE, echo=FALSE}
kable(corr,caption="Table 3.  Correlation between beta cell function outcomes and CGM outcomes.")
```
\

```{r, results='asis',tidy=TRUE, echo=FALSE}
plot(alldata$odi_cpep,alldata$max_sensor,xlab="oDI Cpep",ylab="Max sensor (mg/dL)")
abline(lm(alldata$max_sensor~alldata$odi_cpep))

plot(alldata$odi_cpep,alldata$percent_time_spent_140,xlab="oDI Cpep",ylab="% time >140 mg/dL")
abline(lm(alldata$percent_time_spent_140~alldata$odi_cpep))

plot(alldata$odi_cpep,alldata$mage,xlab="oDI Cpep",ylab="MAGE (mg/dL)")
abline(lm(alldata$mage~alldata$odi_cpep))

```